---
title: "Summary"
output:
  html_document:
    toc: true
    code_folding: "hide"
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(SAJR)
library(visutils)
library(clusterProfiler)
library(Matrix)
library(ggplot2)
castXYtable = function(x,y,i){
  ys = sort(unique(y))
  xs = sort(unique(x))
  m = matrix(NA,nrow=length(xs),ncol=length(ys),dimnames=list(xs,ys))
  x = as.character(x)
  y = as.character(y)
  for(j in 1:length(x))
    m[x[j],y[j]]= i[j]
  m
}

plotMDS = function(xy,celltypes,ct2col,samples,...){
  pch = as.numeric(factor(samples)) %% 25
  plot(xy,col=ct2col[celltypes],pch=pch,xlab='Dim 1',ylab='Dim 2',...)
  for(ct in unique(celltypes)){
    inx = which(celltypes == ct)
    if(length(inx)>1){
      inx = combn(inx,2)
      segments(xy[inx[1,],1],xy[inx[1,],2],xy[inx[2,],1],xy[inx[2,],2],col=ct2col[ct])
    }
  }
  for(ct in unique(celltypes)){
    inx = which(celltypes == ct)
    text(mean(xy[inx,1]),mean(xy[inx,2]),ct)
  }
}

getDitPlotSize = function(ccr){
  size = table(ccr@compareClusterResult$Cluster)
  size = size[size>0]
  size = c(height=sum(pmin(5,size)) * 0.15 + 3,
          width=length(size) * 0.6 + 7)
  size
}

rds.dir = 'rds'
```

# Load data
```{r}
pv = readRDS(paste0(rds.dir,'/all_celltypes_test.rds'))
perct = readRDS(paste0(rds.dir,'/celltypes_marker_test.rds'))
pb = readRDS(paste0(rds.dir,'/pb_as_filtered.rds'))
pbmeta = readRDS(paste0(rds.dir,'/pb_meta_filtered.rds'))
go = readRDS(paste0(rds.dir,'/go.rds'))
ipro = readRDS(paste0(rds.dir,'/ipro.rds'))

gsize = getDitPlotSize(go)
isize = getDitPlotSize(ipro)
```


# Input summary
Number of cells in each celltype * sample
```{r}
ct2sam = castXYtable(pbmeta$celltype,pbmeta$sample_id,pbmeta$ncells)
ct2sam[is.na(ct2sam)] = 0
knitr::kable(ct2sam)
```


# MDS
Segments were first filtered by coverage (>9 reads in at least 20% of sample*celltypes) and psi variability (sd>0.1).
Filtered segments were tested using GLM/LRT using ~celltype model. Segments with fdr < 0.05 and dPSI > 0.3 (between highest and lowest celltype) were considered as significant.
```{r fig.height=18,fig.width=6}
cor_all = cor(pb$ir,use = 'pair') #cor(t(scale(t(pb$ir),scale = FALSE)),use = 'pair')
cor_all[is.na(cor_all)] = 0
mds_all = cmdscale(1-cor_all,k=2)

sgn = pv$celltype_fdr<0.05 & pv$dpsi>0.3
sgn[is.na(sgn)] = FALSE
cor_sgn = cor(pb$ir[sgn,],use = 'pair') # cor(t(scale(t(pb$ir[sgn,]),scale = FALSE)),use = 'pair')
cor_sgn[is.na(cor_sgn)] = 0
mds_sgn = cmdscale(1-cor_sgn,k=2)


ct2col = char2col(pbmeta$celltype)

par(mfrow=c(3,1),tcl=-0.2,mgp=c(1.3,0.3,0),mar=c(3,3,1.5,0),oma=c(0,0,0,0),bty='n',lwd=2)
plotMDS(mds_all,pbmeta$celltype,ct2col,pbmeta$sample_id,main=paste0('Filtered (',length(pb),')'))
plotMDS(mds_sgn,pbmeta$celltype,ct2col,pbmeta$sample_id,main=paste0('Significant (',sum(sgn),')'))
plot.new()
legend(grconvertX(0,'nfc','user'),grconvertY(1,'npc','user'),xpd=NA,bty='n',pch=19,col=ct2col,legend = names(ct2col),title='Celltypes',
       ncol = ceiling(length(ct2col)/15))
```

# Celltype-specific events
Here each celltype was tested against all remaining.
```{r}
sgn = rbind(psi01=apply(perct$fdr<0.05 & abs(perct$dpsi)>0.1 & abs(perct$dpsi) <= 0.5,2,sum,na.rm=TRUE),
            psi05=apply(perct$fdr<0.05 & abs(perct$dpsi)>0.5,2,sum,na.rm=TRUE))
par(mar=c(10,4,1,1))
barplot(sgn,xlab='',ylab='# significant segments',legend.text = c('dPSI > 0.1', 'dPSI > 0.5'),las=2)
```

# GO enrichment
```{r, fig.height=gsize[1],fig.width=gsize[2]}
dotplot(go,label_format=90) + theme(axis.text.x = element_text(angle = 90))
```

# Domain enrichment
```{r, fig.height=isize[1],fig.width=isize[2]}
dotplot(ipro,label_format=90) + theme(axis.text.x = element_text(angle = 90))
```


