---
title: "Summary"
output:
  html_document:
    toc: true
    code_folding: "hide"
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(Matrix)
library(SAJR)
library(clusterProfiler)
library(ggplot2)
library(scsajr)
library(visutils)
rds_dir <- "rds"
```

# Load data
```{r}
pb <- readRDS(paste0(rds_dir, "/pb_as_filtered.rds"))

pv <- pb@metadata$all_celltype_test
perct <- pb@metadata$markers

go <- pb@metadata$go
ipro <- pb@metadata$ipro

gsize <- scsajr::get_dit_plot_size(go)
isize <- scsajr::get_dit_plot_size(ipro)
```


# Input summary
Number of cells in each celltype * sample
```{r}
ct2sam <- scsajr::cast_xy_table(pb$celltype, pb$sample_id, pb$ncells)
ct2sam[is.na(ct2sam)] <- 0
knitr::kable(ct2sam)
```


# MDS
Segments were first filtered by coverage (>9 reads in at least 20% of sample*celltypes) and psi variability (sd>0.1).
Filtered segments were tested using GLM/LRT using ~celltype model. Segments with fdr < 0.05 and dPSI > 0.3 (between highest and lowest celltype) were considered as significant.
```{r fig.height=18,fig.width=6}
psi <- scsajr::calcpsi(pb)
cor_all <- cor(psi, use = "pair") # cor(t(scale(t(pb$ir),scale = FALSE)),use = 'pair')
cor_all[is.na(cor_all)] <- 0
mds_all <- stats::cmdscale(1 - cor_all, k = 2)

sgn <- pv$group_fdr < 0.05 & pv$dpsi > 0.3
sgn[is.na(sgn)] <- FALSE
cor_sgn <- cor(psi[sgn, ], use = "pair") # cor(t(scale(t(pb$ir[sgn,]),scale = FALSE)),use = 'pair')
cor_sgn[is.na(cor_sgn)] <- 0
mds_sgn <- stats::cmdscale(1 - cor_sgn, k = 2)


ct2col <- visutils::char2col(pb$celltype)

par(mfrow = c(3, 1), tcl = -0.2, mgp = c(1.3, 0.3, 0), mar = c(3, 3, 1.5, 0), oma = c(0, 0, 0, 0), bty = "n", lwd = 2)
plotMDS(mds_all, pb$celltype, ct2col, pb$sample_id, main = paste0("Filtered (", length(pb), ")"))
plotMDS(mds_sgn, pb$celltype, ct2col, pb$sample_id, main = paste0("Significant (", sum(sgn), ")"))
plot.new()
legend(graphics::grconvertX(0, "nfc", "user"), graphics::grconvertY(1, "npc", "user"),
  xpd = NA, bty = "n", pch = 19, col = ct2col, legend = names(ct2col), title = "Celltypes",
  ncol = ceiling(length(ct2col) / 15)
)
```

# Celltype-specific events
Here each celltype was tested against all remaining.
```{r}
sgn <- rbind(
  psi01 = apply(perct$fdr < 0.05 & abs(perct$dpsi) > 0.1 & abs(perct$dpsi) <= 0.5, 2, sum, na.rm = TRUE),
  psi05 = apply(perct$fdr < 0.05 & abs(perct$dpsi) > 0.5, 2, sum, na.rm = TRUE)
)
par(mar = c(10, 4, 1, 1))
graphics::barplot(sgn, xlab = "", ylab = "# significant segments", legend.text = c("dPSI > 0.1", "dPSI > 0.5"), las = 2)
```

# GO enrichment
```{r, fig.height=gsize[1],fig.width=gsize[2]}
if (!is.null(go)) {
  clusterProfiler::dotplot(go, label_format = 90) + theme(axis.text.x = element_text(angle = 90))
}
```

# Domain enrichment
```{r, fig.height=isize[1],fig.width=isize[2], eval=!is.null(ipro)}
if (!is.null(ipro)) {
  clusterProfiler::dotplot(ipro, label_format = 90) + theme(axis.text.x = element_text(angle = 90))
}
```

